Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-08-22 manuscript/usses_water_ms.Rmd
updates abstract

  **Methods.** We conducted a five year precipitation manipulation experiment in a sagebrush steppe ecosystem in Idaho, US. We used drought and irrigation treatments (approximately 50% decrease/increase) to investigate whether ecosystem functional response remains consistent at extreme precipitation levels. We recorded data on aboveground net primary productivity (ANPP), species abundance, and soil moisture. We fit a random intercept, random slope model to determine if the relationship between ANPP and soil moisture differed among treatments. We used nondimensional multivariate scaling to quantify community composition over the five years.

  **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP, did not vary across drought, control, and irrigation treatments. ANPP responded positively to availabile soil moisture, but the response was weak and uncertain. There was also no evidence that treatment effects grew over time. Plant community composition was remarkably stable over the course of the experiment and did not differ among treatments.

  **Discussion.** At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is consistent, even at extreme precipitation levels. The similarity of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, but instead reflects the consistent responses of individual species. Such consistency may be due to bet-hedging strategies where species respond weakly to soil moisture to ensure low variance of long-term success.
header-includes: >
  \usepackage{setspace}
  \usepackage{todonotes}
  \usepackage{rotating}
  \usepackage{color, soul}
  \usepackage{sectsty}
  \usepackage{bm,mathrsfs}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
output: 
  rticles::peerj_article:
    number_sections: false
    base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
---

>>>>

  **Methods.** We conducted a five year precipitation manipulation experiment in a sagebrush steppe ecosystem in Idaho, US. We used drought and irrigation treatments (approximately 50% decrease/increase) to investigate whether ecosystem functional response remains consistent at extreme precipitation levels. We recorded data on aboveground net primary productivity (ANPP), species abundance, and soil moisture. We fit a generalized linear mixed effects model to determine if the relationship between ANPP and soil moisture differed among treatments. We used nondimensional multivariate scaling to quantify community composition over the five years.

  **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP was similar among drought and control treatments, but the irrigation treatment had a lower slope than the control treatment. However, ANPP response to availabile soil moisture was weak and uncertain regardless of treatment, with all slopes overlapping zero. There was also no evidence that treatment effects grew over time. Plant community composition was remarkably stable over the course of the experiment and did not differ among treatments.

  **Discussion.** Despite some evidence that ecosystem functional response saturated under extreme wet conditions, the response of ANPP to soil moisture and community compositional changes in this sagebrush steppe system were consistently small. The similarity, or difference, of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, but instead reflects the weak responses of individual species. Such responses may be due to bet-hedging strategies where species respond weakly to even extreme levels of soil moisture to ensure low variance of long-term success.
header-includes: >
  \usepackage{setspace}
  \usepackage{todonotes}
  \usepackage{rotating}
  \usepackage{color, soul}
  \usepackage{sectsty}
  \usepackage{bm,mathrsfs}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
output: 
  rticles::peerj_article:
    number_sections: false
    base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
---

### 2017-08-22 manuscript/usses_water_ms.tex
updates abstract

\textbf{Methods.} We conducted a five year precipitation manipulation
experiment in a sagebrush steppe ecosystem in Idaho, US. We used drought
and irrigation treatments (approximately 50\% decrease/increase) to
investigate whether ecosystem functional response remains consistent at
extreme precipitation levels. We recorded data on aboveground net
primary productivity (ANPP), species abundance, and soil moisture. We
fit a random intercept, random slope model to determine if the
relationship between ANPP and soil moisture differed among treatments.
We used nondimensional multivariate scaling to quantify community
composition over the five years.

\textbf{Results.} Ecosystem functional response, defined as the
relationship between soil moisture and ANPP, did not vary across
drought, control, and irrigation treatments. ANPP responded positively
to availabile soil moisture, but the response was weak and uncertain.
There was also no evidence that treatment effects grew over time. Plant
community composition was remarkably stable over the course of the
experiment and did not differ among treatments.

\textbf{Discussion.} At least in the short-term, ecosystem functional
response and community composition in this sagebrush steppe system is
consistent, even at extreme precipitation levels. The similarity of
ecosystem functional response across treatments was not due to
compensatory shifts at the plant community level, but instead reflects
the consistent responses of individual species. Such consistency may be
due to bet-hedging strategies where species respond weakly to soil
moisture to ensure low variance of long-term success.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

>>>>

\textbf{Methods.} We conducted a five year precipitation manipulation
experiment in a sagebrush steppe ecosystem in Idaho, US. We used drought
and irrigation treatments (approximately 50\% decrease/increase) to
investigate whether ecosystem functional response remains consistent at
extreme precipitation levels. We recorded data on aboveground net
primary productivity (ANPP), species abundance, and soil moisture. We
fit a generalized linear mixed effects model to determine if the
relationship between ANPP and soil moisture differed among treatments.
We used nondimensional multivariate scaling to quantify community
composition over the five years.

\textbf{Results.} Ecosystem functional response, defined as the
relationship between soil moisture and ANPP was similar among drought
and control treatments, but the irrigation treatment had a lower slope
than the control treatment. However, ANPP response to availabile soil
moisture was weak and uncertain regardless of treatment, with all slopes
overlapping zero. There was also no evidence that treatment effects grew
over time. Plant community composition was remarkably stable over the
course of the experiment and did not differ among treatments.

\textbf{Discussion.} Despite some evidence that ecosystem functional
response saturated under extreme wet conditions, the response of ANPP to
soil moisture and community compositional changes in this sagebrush
steppe system were consistently small. The similarity, or difference, of
ecosystem functional response across treatments was not due to
compensatory shifts at the plant community level, but instead reflects
the weak responses of individual species. Such responses may be due to
bet-hedging strategies where species respond weakly to even extreme
levels of soil moisture to ensure low variance of long-term success.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}
